Bibat: Batteries-include Bayesian Analysis Template

### 2024-01-29 paper/main.tex
Update paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Bibat: Batteries-include Bayesian Analysis Template},
  pdfauthor={Teddy Groves},
  pdfkeywords={Bayesian inference,},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Bibat: Batteries-include Bayesian Analysis Template}
\author{Teddy Groves}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Despite their suitability for many scientific problems and the existence
of sound theoretical and computational frameworks, adoption of Bayesian
workflow in science remains a challenge. One reason for the difficulty
is that it is difficult to write software that implements non-trivial
Bayesian workflows. We aimed to address this difficulty by developing
Bibat, a Python package providing an interactive template that targets
Bayesian statistical analysis projects. Bibat trivialises the otherwise
painful process of structuring such a project and integrating together
tools that help with individual parts of a Bayesian workflow, without
compromising flexibility, scalability, interoperability or
reproducibility. Bibat is available on the Python Package index,
documented at \url{https://bibat.readthedocs.io/} and developed at
\url{https://github.com/teddygroves/bibat/}.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

Many areas of science would benefit from replacing traditional
statistical methodologies with Bayesian workflow, as described in, for
example (Gelman et al. 2020; Grinsztajn et al. 2021; Gabry et al. 2019).

There are now software tools that address most individual aspects of a
Bayesian workflow, from fetching and manipulating data to specifying,
computing, storing, inspecting and documenting statistical inferences.
However, writing a software project implementing a Bayesian workflow
remains challenging due partly to the difficulty of orchestrating and
configuring these elements.

Bibat is a Python package that addresses this difficulty by providing an
interactive template for Bayesian workflow projects.

\hypertarget{installation-and-usage}{%
\section{Installation and usage}\label{installation-and-usage}}

Bibat is installed by running the command \texttt{pip\ install\ bibat}
and then used by running the command \texttt{bibat}.

This command triggers an interactive form which prompts the user for
configuration information including project and repository name, author
name, a short description and choices of open source licence options,
documentation formats and whether to include tests and continuous
integration.

Bibat then creates a folder with the chosen repository name in the
current working directory, containing code that implements bibat's
example analysis.

\hypertarget{documentation}{%
\section{Documentation}\label{documentation}}

Bibat is documented at \url{https://github.com/teddygroves/bibat/}. The
documentation website includes instructions for getting started, a
detailed explanation of bibat's concepts, an extended vignette
illustrating intended usage, full description of the python API and
command line interface, instructions for contributing and a section
discussing accessibility considerations.

\hypertarget{software}{%
\section{Software}\label{software}}

Bibat is implemented using Python 3. Its main dependencies are
cookiecutter (Greenfeld et al. 2021), pydantic (Pydantic developers
2022) and click (Click Developers 2022). Bibat is continuously tested to
ensure that it works on the operating systems Linux, macOS and Windows.

Bibat projects use the following tools:

\begin{itemize}
\tightlist
\item
  Python 3 (Van Rossum and Drake 2009) and standard scientific Python
  packages for data manipulation
\item
  Stan (Carpenter et al. 2017) for specifying statistical models and
  performing inferences
\item
  cmdstanpy (Stan Development Team 2022) for interfacing between Python
  and Stan
\item
  arviz (Kumar et al. 2019) for storing and analysing completed
  inferences
\item
  pydantic and pandera (Niels Bantilan 2020) for validation
\item
  make (Stallman and McGrath 1991) for automation.
\item
  Sphinx (Georg Brandl and the Sphinx team 2022) and Quarto (Allaire et
  al. 2022) for documentation.
\end{itemize}

\hypertarget{how-bibat-addresses-specific-bayesian-workflow-difficulties}{%
\section{How bibat addresses specific Bayesian workflow
difficulties}\label{how-bibat-addresses-specific-bayesian-workflow-difficulties}}

This section describes some specific problems that often affect Bayesian
workflow projects and which bibat helps to ameliorate.

\hypertarget{data-modelling}{%
\subsection{Data modelling}\label{data-modelling}}

Any statistical analysis requires a definition of a prepared dataset,
whether implicitly or explicitly. Ideally the definition is explicit and
allows for arbitrarily many prepared datasets, accommodates both tabular
and non-tabular formats and provides functionality like validation and
serialisation.

Bibat achieves this goal by providing abstract models for prepared data,
preconfigured to work well together, integrated into a data preparation
pipeline and able to be easily customised to suit any analysis.

Without a template such as bibat, users must either devise a new data
modelling framework from scratch as part of an analysis, which is
time-consuming, or else to miss out on either the flexibility or
features that bibat's approach provides.

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

Bibat ensures reproducibility by providing a preconfigured makefile with
a target \texttt{analysis} triggering creation of an isolated
environment, installation of dependencies, data preparation, statistical
computation and analysis of results. In this way a bibat analysis can be
reproduced on most platforms using a single command.

@@

\hypertarget{collaboration}{%
\subsection{Collaboration}\label{collaboration}}

Bibat provides a preconfigured test environment, continuous integration,
linting and pre-commit hooks, making it suitable for collaborative
software development. In addition, including documentation as a first
class component of the analysis addresses a common problem in academic
statistics projects where the paper gets out of sync with the code.

\hypertarget{installing-dependencies}{%
\subsection{Installing dependencies}\label{installing-dependencies}}

@@

\hypertarget{fitting-modes}{%
\subsection{Fitting modes}\label{fitting-modes}}

As part of a Bayesian workflow it is often necessary to fit a model and
dataset in different ways. For example, one might perform MCMC sampling
of both the prior and posterior distributions, perform multiple
leave-out-one-fold fits for cross-validation or need to compare MCMC
sampling with an optimisation-based alternative.

Bibat accommodates this ubiquitous scenario by introducing an
abstraction called ``fitting mode'' and a corresponding Pydantic base
class \texttt{FittingMode}, along with and several subclasses for
commonly used cases.

This abstraction allows bibat projects to handle fitting a model and
dataset in different ways appropriately and flexibly. For example, the
provided prior sampling fitting mode creates a Stan input dictionary
with the \texttt{likelihood} data variable set to \texttt{0}, performs
MCMC sampling and writes data to the \texttt{InferenceData} group
\texttt{prior}.

\hypertarget{configure-fitting-concisely.}{%
\subsection{Configure fitting
concisely.}\label{configure-fitting-concisely.}}

Bibat provides a Pydantic class \texttt{InferenceConfiguration}, parsed
from toml files in the \texttt{inferences} directory. By writing a toml
file per inference the user can configure statistical model, dataset,
fitting modes and computational settings (globally or per fitting mode).

This approach appropriately separates configuration from logic and
avoids unnecessary duplication of configuration for runs with the same
model and dataset.

\hypertarget{storing-mcmc-samples}{%
\subsection{Storing MCMC samples}\label{storing-mcmc-samples}}

Bibat provides code that automatically saves its output
\texttt{InferenceData} objects in zarr format. This format distributes
information from the same set of samples into multiple smaller files,
thereby helping to avoid file size limits of online repository hosting
services.

\hypertarget{comparison-with-alternative-software}{%
\section{Comparison with alternative
software}\label{comparison-with-alternative-software}}

Other than bibat, there is currently no interactive template that
specifically targets Bayesian workflow projects. There are some
templates that arguably encompass Bayesian workflow as a special case of
data analysis project, such as cookiecutter-data-science (Driven Data
2022), but these are of limited use compared with a specialised template
due to the many specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but differently from bibat. For example, bambi
(Capretto et al. 2020) and brms (Bürkner 2017) aim to make implementing
Bayesian workflows easier by trivialising the task of specifying a
Bayesian generalised linear model and fitting it to a single tabular
dataset. This approach does not address several difficulties that bibat
does address,

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Allaire_Quarto_2022}{}}%
Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
Christophe Dervieux. 2022. {``Quarto.''}
\url{https://doi.org/10.5281/zenodo.5960048}.

\leavevmode\vadjust pre{\hypertarget{ref-burknerBrmsPackageBayesian2017}{}}%
Bürkner, Paul-Christian. 2017. {``Brms: {An R} Package for {Bayesian}
Multilevel Models Using {Stan}.''} \emph{Journal of Statistical
Software} 80 (1): 1--28. \url{https://doi.org/10.18637/jss.v080.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-capretto2020}{}}%
Capretto, Tomás, Camen Piho, Ravin Kumar, Jacob Westfall, Tal Yarkoni,
and Osvaldo A. Martin. 2020. {``Bambi: {A} Simple Interface for Fitting
{Bayesian} Linear Models in {Python}.''}
\url{https://arxiv.org/abs/2012.10754}.

\leavevmode\vadjust pre{\hypertarget{ref-carpenterStanProbabilisticProgramming2017}{}}%
Carpenter, Bob, Andrew Gelman, Matthew D. Hoffman, Daniel Lee, Ben
Goodrich, Michael Betancourt, Marcus Brubaker, Jiqiang Guo, Peter Li,
and Allen Riddell. 2017. {``Stan: {A Probabilistic Programming
Language}.''} \emph{Journal of Statistical Software} 76 (1): 1--32.
\url{https://doi.org/10.18637/jss.v076.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-clickdevelopersClickPythonComposable2022}{}}%
Click Developers. 2022. {``Click: {Python} Composable Command Line
Interface Toolkit.''} Pallets. \url{https://pypi.org/project/click/}.

\leavevmode\vadjust pre{\hypertarget{ref-drivendataCookiecutterdatascience2022}{}}%
Driven Data. 2022. {``Cookiecutter-Data-Science.''}
\url{https://github.com/drivendata/cookiecutter-data-science/}.

\leavevmode\vadjust pre{\hypertarget{ref-gabryVisualizationBayesianWorkflow2019}{}}%
Gabry, Jonah, Daniel Simpson, Aki Vehtari, Michael Betancourt, and
Andrew Gelman. 2019. {``Visualization in {Bayesian Workflow}.''}
\emph{Journal of the Royal Statistical Society Series A: Statistics in
Society} 182 (2): 389--402. \url{https://doi.org/10.1111/rssa.12378}.

>>>>

\documentclass[sigconf, review]{acmart}

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother

%%
%% This is file `sample-manuscript.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `manuscript')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-manuscript.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}

%% PANDOC PREAMBLE BEGINS

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\definecolor{mypink}{RGB}{219, 48, 122}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
%% PANDOC PREAMBLE ENDS

\setlength{\parindent}{10pt}
\setlength{\parskip}{0pt}

\hypersetup{
  pdftitle={Bibat: Batteries-include Bayesian Analysis Template},
  pdfauthor={Teddy Groves},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={red},
  pdfcreator={LaTeX via pandoc, via quarto}}

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[KDD]{KDD}{Aug 25--29, 2024}{Barcelona, Esp}
\acmPrice{}
\acmISBN{}

%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Bibat: Batteries-include Bayesian Analysis Template}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

  \author{Teddy Groves}
  \orcid{0000-0002-7109-3270}
            \affiliation{%
                  \institution{Danish Technical University}
                                  \city{Kongens Lyngby}
                                  \country{Denmark}
                      }

%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Trovato et al.}
%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Bayesian statistical workflow offers a powerful way to learn from data,
but software software projects that implement complex Bayesian workflows
in practice are unusual, partly due to the difficulty of orchestrating
Bayesian statistical software. Bibat addresses this challenge by
providing a full-featured, scalable Bayesian statistical analysis
project using an interactive template. Bibat is available on the Python
Package index, documented at \url{https://bibat.readthedocs.io/} and
developed at \url{https://github.com/teddygroves/bibat/}. This paper
explains the motivation for bibat, briefly describes intended usage,
discusses key design choices, and reviews several examples of bibat's
use in scientific applications.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Python, Bayesian workflow, Methodology}

\begin{teaserfigure}
  \includegraphics[width=\textwidth]{img/workflow.png}
  \caption{Schematic representation of a Bayesian workflow implemented
using bibat. The author inspects their analysis's results, edits files
corresponding to the boxes on the left, runs the command
\texttt{make\ analysis}, then repeats. Note that this workflow is
modular, accommodates plurality and results in a final analysis that can
be fully reproduced using a single command.}
  \Description{DESCRIPTION}
\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\setlength{\parskip}{-0.1pt}

\section{Introduction: the problem of orchestrating Bayesian workflow
software}\label{introduction-the-problem-of-orchestrating-bayesian-workflow-software}

The idea that Bayesian statistical analysis comprises not just
inference, but also specific approaches to related activities like data
preparation, model design, diagnosis, debugging and criticism, dates at
least to \citet{boxBayesianInferenceStatistical1992}. There is
increasing scholarly recognition of the need for a holistic view of
``Bayesian workflow''
\citep{gelmanBayesianWorkflow2020, grinsztajnBayesianWorkflowDisease2021, gabryVisualizationBayesianWorkflow2019}.
Relatedly, software tools now exist that address most individual aspects
of a Bayesian workflow: see \citet{strumbeljPresentFutureSoftware} for a
review of the state of the art.

Unfortunately, currently available tools typically address one, or at
most a handful, of Bayesian workflow activities; it is left to the
individual project team to orchestrate the components. Writing software
that performs this orchestration can be time-consuming and tricky,
especially in the common scenario where it is not initially clear how
many, or what kind of, statistical models, datasets, data manipulations
or investigations an analysis will require.

Bibat addresses this difficulty by providing a full-featured,
high-quality Bayesian workflow project that can be extended to implement
a wide range of statistical analyses.

\section{How bibat works}\label{how-bibat-works}

\subsection{Installation and usage}\label{installation-and-usage}

Bibat can be installed on the Windows, linux or macos command line on by
running the command \texttt{pip\ install\ bibat} and then used by
running the command \texttt{bibat}. This command triggers an interactive
form which prompts the user to select a range of customisation options.
Bibat then creates a new directory containing code that implements an
example analysis, with customisations reflecting the user's choices.
This analysis works immediately, and can be reproduced with the single
command \texttt{make\ analysis} without the need for any further action
by the user: in this sense bibat comes with ``batteries included''.

\subsection{Documentation}\label{documentation}

Bibat is documented at \url{https://github.com/teddygroves/bibat/}. The
documentation website includes instructions for getting started, a
detailed explanation of bibat's concepts and an extended vignette
illustrating how to implement a complex statistical analysis starting
from bibat's example analysis usage. In addition, the documentation site
contains a full description of bibat's python API and command line
interface, instructions for contributing and a section discussing
accessibility considerations.

\section{Design choices}\label{design-choices}

Bibat's design was informed by these considerations.

\subsection{Accommodate a wide range of statistical
analyses}\label{accommodate-a-wide-range-of-statistical-analyses}

Bibat projects explicitly allow for plurality at the level of input
datasets, prepared datasets, statistical models, fitting modes,
computation methods and analyses. In addition, bibat ensures that there
are minimal restrictions on the kind of components: for example,
datasets need not be singular or tabular, and statistical models need
not be representable in formula syntax.

Thanks to these accommodations a project team using bibat should
typically not need to foresee the ultimate requirements of their
analysis before starting the project.

\subsection{Encourage reproducibility}\label{encourage-reproducibility}

Bibat provides a preconfigured makefile with a target \texttt{analysis}
triggering creation of an isolated environment, installation of
dependencies, data preparation, statistical computation and analysis of
results. In this way a bibat analysis can be reproduced on most
platforms using a single command.

@@

\subsection{Use widely-adopted, open source and actively developed
tools}\label{use-widely-adopted-open-source-and-actively-developed-tools}

Bibat projects are is written in modern Python and uses pydantic
\citet{pydanticdevelopersPydantic2022} and pandera
\citep{niels_bantilan-proc-scipy-2020} for data modelling, Stan
\citep{carpenterStanProbabilisticProgramming2017} for statistical
inference, cmdstanpy \citep{standevelopmentteamCmdStanPy2022} for
python-Stan interface, arviz \citep{kumarArviZUnifiedLibrary2019} for
storing and analyzing inferences and sphinx
\citep{georgbrandlandthesphinxteamSphinx2022} and quarto
\citep{Allaire_Quarto_2022} for documentation.

Bibat itself is also written in modern Python and uses the popular tools
cookiecutter \citep{greenfeldCookiecutter2021}, pydantic and click
\citep{clickdevelopersClickPythonComposable2022}.

@@

\subsection{Implement community standards and best practices for
collaborative software
development}\label{implement-community-standards-and-best-practices-for-collaborative-software-development}

Bibat projects include a preconfigured test environment, continuous
integration, linting and pre-commit hooks, making them suitable for
collaborative software development. In addition, including documentation
as a first class component of the analysis addresses a common problem in
academic statistics projects where the paper gets out of sync with the
code.

Bibat itself is continuously tested to ensure that it works on the
operating systems Linux, macOS and Windows. Bibat's continuous
integration runs a test suite as well as an end-to-end functional test
on all supported Python versions.

Bibat is part of the PyOpenSci ecosystem, allowing for community help
with maintenance as well as peer review for code and documentation
quality, usability and accessibility. The PyOpenSci peer review for
bibat can be found here:
\url{https://github.com/pyOpenSci/software-submission/issues/83}

\section{Challenges}\label{challenges}

This section describes some specific challenges that often affect
Bayesian workflow projects and how bibat addresses them.

\subsection{Address complexity using a modular, file based
approach}\label{address-complexity-using-a-modular-file-based-approach}

As discussed in \citet{gelmanBayesianWorkflow2020}, Bayesian workflows
are complicated, featuring plurality, cyclicity and complexity at many
levels. Bibat accommodates this by separating non-interacting analysis
components into modules and by serialising data to files wherever
possible. Prepared datasets, statistical models, inference
configurations, inference results, plots and analyses all have file
representations. Fitting modes, data manipulations and data models are
modularised in code through the use of appropriately structured data
classes and functions.

Thanks to this approach it is possible to perform small sub analyses
individually and to iteratively add components without needing to
consider everything at once.

\subsection{Fitting modes}\label{fitting-modes}

As part of a Bayesian workflow it can be necessary to fit a model and
dataset in different ways. For example, one might perform MCMC sampling
of both the prior and posterior distributions, perform multiple
leave-out-one-fold fits for cross-validation or need to compare MCMC
sampling with an optimisation-based alternative.

Bibat accommodates this need by introducing an abstraction called
``fitting mode''. This abstraction allows bibat projects to handle
fitting a model and dataset in different ways appropriately and
flexibly. For example, the provided prior sampling fitting mode creates
a Stan input dictionary with the \texttt{likelihood} data variable set
to \texttt{0}, performs MCMC sampling and writes data to the
\texttt{InferenceData} group \texttt{prior}. Bibat provides fitting
modes corresponding for prior sampling, posterior sampling and k-fold
posterior sampling. Users can easily add additional fitting modes by
copying these examples.

\section{Post launch use}\label{post-launch-use}

Bibat has been used for several real statistical analyses.

\citet{grovesteddyDgfreg2023} used bibat to compare a Bayesian and two
non-Bayesian approaches to modelling a biochemical thermodynamics
dataset. Bibat facilitated this analysis even though it was not very
large by providing the fitting mode abstraction, which was very useful
for comparing the different methods.

In \citet{grovesBaseball2022}, Bibat was used to implement a sports
analysis involving two datasets, two models and four inferences,
demonstrating that the generalised Pareto distribution can be used to
describe hitting ability in baseball. This analysis is now included in
bibat as an illustration, along with an accompanying tutorial.

\begin{figure}

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{img/baseball.png}

}

\caption{\label{fig-baseball}A graphical posterior predictive check
produced as part of a bibat analysis that fit two statistical models to
two datasets of baseball data. The coloured lines show each model's
posterior predictive distributions and the black dots show the two
observed datasets. See
\url{https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball}
for the full analysis.}

\end{figure}%

In \citet{grovesteddySphincter2024}, bibat was used to implement an
analysis of cerebrovascular data from mice, involving two raw datasets,
6 prepared datasets, 15 models and 15 inferences. In this case bibat's
modular design made it straightforward to iteratively add data
transformations and models while maintaining reproducibility.

These cases illustrate that bibat can be useful in a variety of real
Bayesian workflows, with different sizes, subject matters and emphases.

Bibat has a growing user community, with 16 GitHub stars at the time of
writing.

\section{Comparison with alternative
software}\label{comparison-with-alternative-software}

Other than bibat, there is currently no interactive template that
specifically targets Bayesian workflow projects. There are some
templates that arguably encompass Bayesian workflow as a special case of
data analysis project, such as cookiecutter-data-science
\citep{drivendataCookiecutterdatascience2022}, but these are of limited
use compared with a specialised template due to the many specificities
of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but using a different approach from bibat's. For
example, bambi \citep{capretto2020} and brms
\citep{burknerBrmsPackageBayesian2017} aim to make implementing Bayesian
workflows easier by providing ergonomic ways to specify and fit Bayesian
regression models to tabular datasets. Bibat is complementary with these
packages, as it targets use cases that they do not support, such as
analyses where complex datasets or custom models might be required.

\section{Statements and Declarations}\label{statements-and-declarations}

The authors have no financial or non-financial interests that are
directly or indirectly related to the work submitted for publication.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography.bib}

%% begin pandoc before-bib
%% end pandoc before-bib
%% begin pandoc biblio
%% end pandoc biblio
%% begin pandoc include-after
%% end pandoc include-after
%% begin pandoc after-body
%% end pandoc after-body
